---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(tidyr)
fn="BirdStrikesData-V2.csv"
df_raw<-read.csv(file=fn,header=T,stringsAsFactors = F,na.strings=c("","NA","N/A"))
```

```{r}
head(df_raw,5)
```

```{r}
date_sep <- substring(df_raw$flight_date,1,10)
date_format <- gsub("-","/",date_sep)
df_raw$flight_date <- as.Date(date_format, '%m/%d/%Y')
head(df_raw,5)
```

```{r}
df_raw <- df_raw %>% replace_na(list(aircraft="Unknown",airport="Unknown",flight_date="Unknown",airline="Unknown",pilot_warned_flag="Unknown",aircraft="Unknown",airport="Unknown",flight_date="Unknown",airline="Unknown",pilot_warned_flag="Unknown",altitude_ft="Unknown",origin="Unknown"))
```

```{r}
colSums(is.na(df_raw))
```


```{r}
df_raw$pilot_warned_flag[df_raw$pilot_warned_flag== "Y"] <- TRUE
df_raw$pilot_warned_flag[df_raw$pilot_warned_flag== "N"] <- FALSE
```

```{r}
library(dplyr)
df_raw = filter(df_raw, airline != "MILITARY")
```

```{r}
drop <- c("impact","wildlife_struck","damage","remains_collected_flag", "Remarks","wildlife_size","species", "heavy_flag")
  
df_new = df_raw[,!(names(df_raw) %in% drop)]
```


```{r}
unique(df_new$flight_phase)
```

```{r}
df_new$flight_phase<-sub("Climb|Take-off run","Take Off",df_new$flight_phase)
df_new$flight_phase<-sub("Landing Roll|Descent","Landing",df_new$flight_phase)
df_new$flight_phase<-sub("Approach","In Flight",df_new$flight_phase)
df_new$flight_phase<-sub("Taxi|Parked","Unknown",df_new$flight_phase)
df_new$flight_phase[df_new$flight_phase==''] <- 'Unknown'
```

```{r}
df_new$airline<-gsub('[*]', '', df_new$airline)
head(df_new,5)
```

```{r}
#1. library
#install.packages('RMySQL')
library(RMySQL)
library(DBI)
#2. Setting 
db_user <- 'admin'
db_password <- 'Amazon3443'
db_name <- 'SandboxDB'
db_host <- 'cs5200sid.c8u2oysrjplq.us-east-2.rds.amazonaws.com' #aws host
db_port <- 3306

#3 Read data from db
mydb <- dbConnect(MySQL(), user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
```

#Creating all the basic tables.
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS STATE(
STATEID INT AUTO_INCREMENT PRIMARY KEY,
STATENAME VARCHAR(80) UNIQUE NOT NULL,
STATECODE VARCHAR(10)
)
```

#AIRCRAFT look up table.
```{sql connection=mydb}
create table if not exists AIRCRAFT(
ACID INT AUTO_INCREMENT PRIMARY KEY,
MODEL VARCHAR(15) NOT NULL UNIQUE
)
```

#FLIGHTPHASE lookup table.
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS FLIGHTPHASE(
FID INT PRIMARY KEY AUTO_INCREMENT,
FLIGHTPHASE VARCHAR(20) NOT NULL UNIQUE
);
```

#CONDITIONS TABLE
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS CONDITIONS(
CID INT AUTO_INCREMENT PRIMARY KEY,
SKY_CONDITION VARCHAR(100) NOT NULL UNIQUE,
EXPLANATION TEXT
);
```

#AIRPORTS TABLE
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS AIRPORTS(
STATE VARCHAR(80),
AIRPORTNAME VARCHAR(180) NOT NULL UNIQUE,
AIRPORTCODE VARCHAR(10),
AID INT PRIMARY KEY AUTO_INCREMENT,
 FOREIGN KEY(STATE) REFERENCES STATE(STATENAME)
);

```

#AIRLINE TABLE 
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS AIRLINE(
AIRLINEID INT AUTO_INCREMENT PRIMARY KEY,
AIRLINENAME VARCHAR(190) NOT NULL UNIQUE
)
```

#INCIDENTS TABLE
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS INCIDENTS(
RID INT PRIMARY KEY,
DATE DATE NOT NULL,
ORIGIN VARCHAR(80),
FOREIGN KEY(ORIGIN) REFERENCES AIRPORTS(STATE),
AIRLINE VARCHAR(190),
FOREIGN KEY(AIRLINE) REFERENCES AIRLINE(AIRLINENAME),
AIRCRAFT VARCHAR(15),
FOREIGN KEY(AIRCRAFT) REFERENCES AIRCRAFT(MODEL),
FLIGHTPHASE VARCHAR(20),
FOREIGN KEY(FLIGHTPHASE) REFERENCES FLIGHTPHASE(FLIGHTPHASE),
ALTITUDE DOUBLE,
CONDITIONS VARCHAR(100),
FOREIGN KEY(CONDITIONS) REFERENCES CONDITIONS(SKY_CONDITION),
WARNING BOOL
)
```

#Connecting to mysql rds on aws.
```{r}
#1. library
#install.packages('RMySQL')
library(RMySQL)

#2. Setting 
db_user <- 'admin'
db_password <- 'Amazon3443'
db_name <- 'SandboxDB'
db_host <- 'cs5200sid.c8u2oysrjplq.us-east-2.rds.amazonaws.com' #aws host
db_port <- 3306

#3 Read data from db
mydb <- dbConnect(MySQL(), user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
```

```{r}
stateVals <- distinct(df_new_without_empty_records,origin)
```

#Adding state values to state table.
#Let's first inspect the state table.

```{sql connection=mydb}
desc STATE;
```

#So, state code can be null, and we have kept stateid as autoincrement. Let's add all the names 
```{r}
stateVals <- stateVals[order(stateVals$origin),]
n.states <- ncol(stateVals)
for (r in stateVals){
  rs <- dbSendStatement(conn = mydb, 
                        statement = paste0("insert into STATE(STATENAME) values (\'",r,"\')"))
  dbClearResult(rs)
}

```

```{sql connection=mydb}
SELECT * FROM STATE;
```

```{r}
aircraftVals <- distinct(df_new_without_empty_records,model)
```

#Adding AIRCRAFT values to AIRCRAFT table.
#Let's first inspect the AIRCRAFT table.

```{sql connection=mydb}
desc AIRCRAFT ;
```

#So, aircraft model cannot be null, and we have kept acid as autoincrement. 
```{r}
aircraftVals <- aircraftVals[order(aircraftVals$model),]
n.aircrafts <- ncol(aircraftVals)
for (r in aircraftVals){
  rs <- dbSendStatement(conn = mydb, 
                        statement = paste0("insert into AIRCRAFT(MODEL) values (\'",r,"\')"))
  dbClearResult(rs)
}
```

```{sql connection=mydb}
SELECT * FROM AIRCRAFT;
```

```{r}
flight <- distinct(df_new_without_empty_records,flight_phase)
```

#Adding FLIGHTPHASE values to FLIGHTPHASE table.
#Let's first inspect the FLIGHTPHASE table.

```{sql connection=mydb}
desc FLIGHTPHASE ;
```

#So, flightphase cannot be null, and we have kept fid as autoincrement.
```{r}
flight <- flight[order(flight$flight_phase),]
n.flight <- ncol(flight)
for (r in flight){
  rs <- dbSendStatement(conn = mydb,
                        statement = paste0("insert into FLIGHTPHASE(FLIGHTPHASE) values (\'",r,"\')"))
  dbClearResult(rs)
}
```

```{sql connection=mydb}
SELECT * FROM FLIGHTPHASE;
```
#conditions values
```{r}
conditions <- distinct(df_new_without_empty_records,sky_conditions)
```

#Adding flightphase values to flightphase table.
#Let's first inspect the flightphase table.

```{sql connection=mydb}
desc CONDITIONS;
```

#So, CONDITIONS cannot be null, and we have kept Cid as autoincrement.
```{r}
conditions <- conditions[order(conditions$sky_conditions),]
n.conditions <- ncol(conditions)
for (r in conditions){
  rs <- dbSendStatement(conn = mydb,
                        statement = paste0("insert into CONDITIONS(SKY_CONDITION) values (\'",r,"\')"))
  dbClearResult(rs)
}
```

```{sql connection=mydb}
SELECT * FROM CONDITIONS;
```













